---
bibliography: references.bib
csl: references.csl
output:
  html_document:
    theme: null
    df_print: kable
params:
  reports_dir: /sbidata/server/daniel/src/back_end/reports/
  project_dir: /sbidata/server/daniel/userdat/example
  db_dir: /sbidata/server/daniel/db/
---

```{r, include=FALSE}
source("defaults.R")

project <-
  paste0(params$project_dir, "/input/project.json") %>%
  jsonlite::read_json(simplifyVector = TRUE)

selected_params <- get_selected_params(project)

analysis_params <- project$params$analysis_params[[selected_params$analysis]]
analysis_dir <- paste0(params$project_dir, "/analysis/", stringr::str_replace_all(selected_params$analysis, " ", "_"), "/")
features_dir <- paste0(params$project_dir, "/features/", stringr::str_replace_all(selected_params$features, " ", "_"), "/")
correlation_dir <- paste0(analysis_dir, "correlation/")
pval_max <- analysis_params$max_pvalue
cor_min <- analysis_params$correlation_min_abs_cor

features_meta_tbl <-
  features_dir %>%
  paste0("features.meta.csv") %>%
  readr::read_csv()

cor_groups <-
  paste0(correlation_dir, "/groups.txt") %>%
  read_lines() %>%
  # correlation result must be available
  purrr::keep(~ paste0(correlation_dir, .x, "/results.csv") %>% file.exists())

cor_groups_discarded <-
  paste0(correlation_dir, "/groups.txt") %>%
  read_lines() %>%
  purrr::discard(~ paste0(correlation_dir, .x, "/results.csv") %>% file.exists())

cor_tbl <-
  tibble(group = cor_groups) %>%
  mutate(
    cor_results_path = paste0(correlation_dir, group, "/results.csv"),
    cor_results_tbl = map2(cor_results_path, group, ~ readr::read_csv(.x) %>% dplyr::mutate(group = .y))
  ) %>%
  pull(cor_results_tbl) %>%
  bind_rows() %>%
  # filter by p value if provided. This is needed to pass banocc results
  dplyr::filter(if (p_value %>% is.na() %>% all()) TRUE else p_value <= pval_max) %>%
  dplyr::filter(abs(cor) >= cor_min)

```

```{r, include=FALSE, eval=cor_groups %>% length() > 0}
groups_cor_tbl <- danielLib::get_groups_cor_tbl(correlation_dir = correlation_dir, nodes_tbl = features_meta_tbl, max_p_value = pval_max, min_abs_cor = cor_min)

node_topology_tbl <- danielLib::get_node_topology_tbl(groups_cor_tbl, node_topology_properties = c("Betweeness centrality", "Closeness centrality", "Node degree"))

correlated_features <-
  cor_tbl %>%
  dplyr::select(matches("^feature")) %>%
  tidyr::gather(k, v) %>%
  dplyr::pull(v) %>%
  unique()

has_many_cor_groups <- groups_cor_tbl %>% dim() %>% ifelse(is.null(.), 0, .) %>% pluck(1) > 1
```

<script>MathJax.Hub.Queue(["Typeset", MathJax.Hub]);</script>

# Correlation Networks

Correlation is used to assess co-abundant features.

```{rmd, eval=cor_groups %>% length() == 0}
## Results  

- No correlation network was calculated
- Possible reasons:
  - Lack of samples or features in all sample groups
  `r {ifelse(analysis_params$correlation_method == "banocc", "- BAnOCC models did not reach convergence", "")}`
```
```{r, eval=cor_groups %>% length() == 0}
knitr::knit_exit()
```

## Methods

```{rmd, eval = analysis_params$correlation_grouping != "all"}
- Correlations were calculated separately using sample grouping `r {analysis_params$correlation_grouping}`
- Only groups having at least `r {analysis_params$correlation_min_samples}` samples and `r {analysis_params$correlation_min_features}` features were considered
- Sample groups matched these criteria: `r {cor_groups %>% paste0(collapse = ", ")}`
```
```{rmd, eval = analysis_params$correlation_grouping == "all"}
- Correlations were calculated using all samples combined
```
- Only correlations with absolute correlation coefficient $abs(r) \ge `r {cor_min}`$ were considered
```{rmd, eval = analysis_params$correlation_method != "banocc"}
- Only correlations with $p \le `r {pval_max}`$ were considered
```
```{rmd, eval = analysis_params$correlation_method == "sparcc"}
- FastSpar implementation of SparCC was used to calculate correlations between features [@pmid30169561; @pmid23028285]
- This method is optimized for sparse data
- Significance was evaluated using `r {analysis_params$sparcc_repetitions}` bootstraps and iterations
```
```{rmd, eval = analysis_params$correlation_method == "banocc"}
- BAnOCC was used to calculate correlations between features [@pmid29140991]
- This method is optimized for compositional data
- Raw feature counts were normalized using total sum scaling (cumulative abundance of each sample is 1)
- Parameters used for the Bayesian model: `r {analysis_params$banocc_chains}` chains, `r {analysis_params$banocc_iters}` iterations of which `r {analysis_params$banocc_warmup}` were used for warm-up, and a minimum confidence level of $\alpha \ge `r {analysis_params$banocc_alpha}`$
```

## Results
```{rmd, eval=cor_tbl %>% nrow() == 0}
- No correlation was significant in any sample group

## References
```
```{r, eval=cor_tbl %>% nrow() == 0}
knitr::knit_exit()
```
- Totaling `r {correlated_features %>% length()}` features were significantly correlated to at lest one other feature ($p \leq `r {pval_max}`$)
```{rmd, eval=cor_groups_discarded %>% length() > 0}
- Sample groups in which calculating correlation network failed: `r {cor_groups_discarded %>% paste(collapse = ", ")}`
```
```{rmd, eval = correlated_features %>% length() > 0}
- These are namely: `r {correlated_features %>% paste(collapse = ', ')}`
```
```{rmd, eval = analysis_params$correlation_method != "banocc"}
- This resulted in `r {cor_tbl %>% dim() %>% magrittr::extract(1)}` significant correlations ($p \leq `r {pval_max}`$)
```
```{rmd, eval = analysis_params$correlation_method == "banocc"}
- This resulted in `r {cor_tbl %>% dim() %>% magrittr::extract(1)}` significant correlations ($\alpha \geq `r {analysis_params$banocc_alpha}`$)
```
```{r, eval = correlated_features %>% length() > 0}
cor_tbl %>%
  dplyr::select(
    group,
    `Feature A` = feature_a, `Feature B` = feature_b, `Correlation coeff` = cor, `p value` = p_value, `q value` = q_value
  ) %>%
  kable(caption = "Significant correlations") %>%
  kableExtra::collapse_rows(1) %>%
  kableExtra::scroll_box(height = "500px")
```

```{rmd, eval = groups_cor_tbl %>% dim() %>% pluck(1) %>% ifelse(is.null(.), 0, .) > 1}
- Closeness centrality, betweeness centrality and node degree were calculated to asses node topology
- Distribution of these metrics were compared between the sample group using Kruskal-Wallis test and Dunn's test
```

```{r, eval = ! is.null(groups_cor_tbl)}
cor_net_plt <-
  groups_cor_tbl %>%
  pull(plt) %>%
  ggpubr::ggarrange(plotlist = ., common.legend = TRUE, legend = "right")

plot_girafe(cor_net_plt)
```

```{r, eval = has_many_cor_groups}
node_topology_stat_tbl <-
  node_topology_tbl %>%
  group_by(key) %>%
    do(
      quietly({
        dunn.test::dunn.test(x = .[["value"]], g = .[["correlation_group"]], kw = FALSE)
      }) %>%
        as.list() %>%
        tibble::as_tibble() %>%
        tidyr::separate(comparisons, into = c("group1", "group2"), sep = " - ") %>%
        dplyr::mutate(label = P.adjusted %>% significance_label())
  ) %>%
  dplyr::filter(P.adjusted <= pval_max)

node_topology_stat_tbl %>%
  dplyr::select(Metric = key, `Group A` = group1, `Group B` = group2, `q value` = P.adjusted) %>%
  kable(caption = "Significant topological metrics") %>%
  kableExtra::collapse_rows(1)
```

```{rmd, eval = has_many_cor_groups}
### Network topology
- network topology differed significantly in `r {node_topology_stat_tbl %>% mutate(comp = paste0(group1, " vs. ", group2)) %>% pull(comp) %>% unique() %>% paste(collapse = ", ")}` (Kruskal-Wallis test, Dunn's test, $p \leq `r {pval_max}`$)
```
  
```{r}
topology_plt <- danielLib::plot_correlation_topology(node_topology_tbl)

plot_girafe(topology_plt)
```

```{r, eval = analysis_params$correlation_method == "banocc"}
banocc_envs <-
  cor_groups %>%
  map_chr(~ paste0(correlation_dir, .x, "/environment.RData")) %>%
  map( ~ {
    env <- new.env()
    load(.x, envir = env)
    as.list(env)
  })

fits_names <-
  banocc_envs %>%
  purrr::map_chr(~ .x %>% purrr::pluck("out_dir") %>% basename())

fits <-
  banocc_envs %>%
  purrr::map(~ .x %>% purrr::pluck("b_fit", "Fit")) %>%
  magrittr::set_names(fits_names)
  
fits_summary_tbl <-
  banocc_envs %>%
  map(~ .x %>% pluck("b_stat_tbl")) %>%
  magrittr::set_names(fits_names) %>%
  tibble::enframe() %>%
  unnest(cols = c(value))

params_converged <- fits_summary_tbl %>% dplyr::filter(Rhat < 1.05) %>% nrow()
params_total <- fits_summary_tbl %>% nrow()
```

```{rmd, eval = analysis_params$correlation_method == "banocc"}
### Model fit

- BAnOCC model was fitted using rstan engine
- Convergence was achieved for `r {params_converged}` out of `r {params_total}` parameters ($\hat R < 1.05$)
```

```{r, eval = analysis_params$correlation_method == "banocc", fig.cap="Model convergence. End of warmup is indicated by black vertical lines."}
fits %>%
  danielLib::plot_stanfit() %>%
  plot_girafe()
```

```{r, eval = analysis_params$correlation_method == "banocc"}
fits_summary_tbl %>%
  kable(caption = "Summary of model fit. Estimated parameters and derived quantities using the posterior draws.") %>%
  kableExtra::collapse_rows(1) %>%
  kableExtra::scroll_box(height = "500px")
```

## Notes
- Correlation is neither sufficient nor required to derive causality from it
```{rmd, eval = analysis_params$correlation_method == "spearman"}
- Spurious correlations
  - Reasons
    - Spearman\'s method is prone to spurious correlations
  - Implications
    - A very high false positive rate is expected
  - Suggestions
    - Apply a high prevalence filter threshold to reduce sparseness
    - Use other correlations aware of sparseness (e.g. SparCC)
```

## References
