---
bibliography: references.bib
csl: references.csl
output:
  html_document:
    theme: null
    df_print: kable
params:
  reports_dir: /sbidata/server/daniel/latest/repo/back_end/reports/
  project_dir: /sbidata/server/daniel/latest/userdat/test_single/
  db_dir: /sbidata/server/daniel/latest/db/
---

```{r, include=FALSE}
source("defaults.R")

library(vegan)
library(phyloseq)
library(ComplexHeatmap)

project <-
  paste0(params$project_dir, "/input/project.json") %>%
  jsonlite::read_json(simplifyVector = TRUE)

selected_params <- danielLib::get_selected_params(project)

features_params <- project$params$features_params[[selected_params$features]]
features_dir <- paste0(params$project_dir, "/features/", stringr::str_replace_all(selected_params$features, " ", "_"), "/")

samples_tbl <-
  paste0(params$project_dir, "/input/samples.csv") %>%
  read_csv_guess_coltypes()

phylo_db <-
  paste0(params$db_dir, "/phylogenies/db.csv") %>%
  readr::read_csv() %>%
  dplyr::filter(id == features_params$phylo_database) %>%
  as.list()

available <-
  paste0(features_dir, "features.raw.csv") %>%
  file.exists()
```

```{r, eval=available, include = FALSE}
features_raw_tbl <-
  paste0(features_dir, "features.raw.csv") %>%
  readr::read_csv()

features_tbl <-
  paste0(features_dir, "features.csv") %>%
  readr::read_csv()

features <-
  features_tbl %>%
  colnames() %>%
  setdiff("sample_id")

features_phy <-
  paste0(features_dir, "phyloseq.rds") %>%
  readr::read_rds()

alpha_div_tbl <- try(danielLib::get_alpha_diversity_all(features_raw_tbl), silent = TRUE)
alpha_div_metrics <- try(
  alpha_div_tbl %>%
  colnames() %>%
  setdiff("sample_id")
)
```

# Features

Feature generation is the process of pooling denoised sequences at a given taxonomic rank.
Furthermore, abundance profiles were normalized to make them comparable between differentially sequenced samples.

## Methods
- Abundances of denoised sequences were pooled to `r {features_params$taxonomic_rank}` rank
- Phylogeny used: `r {phylo_db$name}`: `r {phylo_db$description}`
- Normalization method used: `r {features_params$normalization_method}`
- Total sum scaling (TSS) was used for ordination
- A feature is considered as prevalent if it's abundance is at least `r {features_params$min_abundance}`%
- Only features more prevalent than `r {features_params$min_prevalence}`% of all samples were kept for downstream analysis
- Strategy `r {features_params$unknown_strategy}` was applied to deal with unassigned sequences

## Results

```{rmd, eval = !available}
- Feature generation failed
```

```{r, eval = !available}
knitr::knit_exit()
```

- Totaling `r {length(features)}` features and `r {samples_tbl %>% nrow()}` samples were generated


```{rmd, eval = 'data.frame' %in% class(alpha_div_tbl)}
- Alpha diversity on feature level (`r {features_params$taxonomic_rank}`)
  - Shannon ranged `r {alpha_div_tbl$Shannon %>% summary_text(num_fmt = ".2f")}`
  - Chao1 ranged `r {alpha_div_tbl$Chao1 %>% summary_text(num_fmt = ".2f")}`
  - Inverse Simpson index ranged `r {alpha_div_tbl$InvSimpson %>% summary_text(num_fmt = ".2f")}`
```

```{r, fig.cap="Feature alpha diversity"}
try(
 danielLib::plot_alpha_diversity(abundance_tbl = features_raw_tbl, samples_tbl = samples_tbl) %>% ggiraph::girafe(ggobj = .)
)
```

```{r, fig.cap = "Abundance profile"}
try(
 danielLib::plot_abundance_clustermap(
  features_tbl = features_tbl,
  normalization_method = features_params$normalization_method,
  taxonomic_rank = features_params$taxonomic_rank,
  clustering_distance = "euclidean"
)
)
```

```{r}
features_ord_plt <- try(danielLib::plot_beta_diversity(features_phy), silent = TRUE)
```

```{r, fig.cap = "Ordination: Between groups beta diversity", eval = "ggplot" %in% class(features_ord_plt)}
ggiraph::girafe(ggobj = features_ord_plt)
```

```{rmd, eval= class(features_ord_plt) == "try-error"}
- Ordination plot could not be generated: `r {features_ord_plt[1]}`
```

