---
bibliography: references.bib
csl: references.csl
output:
  html_document:
    theme: null
    df_print: kable   
params:
  reports_dir: /sbidata/server/daniel/src/back_end/reports/
  project_dir: /sbidata/server/daniel/userdat/test_example/
  db_dir: /sbidata/server/daniel/db/
---

```{r, include=FALSE}
source("defaults.R")

project <-
  paste0(params$project_dir, "/input/project.json") %>%
  jsonlite::read_json(simplifyVector = TRUE)

input_params <- project$input

input_dir <- paste0(params$project_dir, "/input/")

samples_tbl <-
  base::paste0(params$project_dir, "/input/samples.csv") %>%
  danielLib::read_csv_guess_coltypes() %>%
  dplyr::group_by(sample_id)

reads_checksums_tbl <-
  base::paste0(params$project_dir, "/input/reads/checksums.csv") %>%
  readr::read_csv()
```

# Input

- Uploaded samples: `r {input_params$local_samples}`
- Integrated existing projects
  - Studies: `r {input_params$projects$projects}`
  - Query used to filter these samples: `r {input_params$projects$filter_query}`
  - Samples: `r {input_params$projects$samples}`
- Integrated NCBI SRA run ids: `r {input_params$sra_ids}`
- Samples selected for demultiplexing: `r {input_params$muxed_samples}`
- Only samples occurring in the samples meta data table were considered for downstream analysis

```{r}
samples_tbl %>%
  kable(caption = "Sample meta data") %>%
  kable_styling() %>%
  scroll_box(height = "500px")
```

```{r}
samples_tbl %>%
  dplyr::ungroup() %>%
  dplyr::select_if(is.factor) %>%
  tidyr::gather(k, v) %>%
  ggplot2::ggplot(aes(v)) +
  ggplot2::geom_bar() +
  ggplot2::coord_flip() +
  ggplot2::facet_wrap(~k, scales = "free_y") +
  ggplot2::labs(x = "", y = "Samples count")
```

```{r, fig.cap="MD5 checksums of raw read files"}
reads_checksums_tbl %>%
  dplyr::mutate(
    file_prefix = file %>% stringr::str_remove("_(1|2)\\.raw\\.fq\\.gz$"),
    mate = file %>% stringr::str_extract("(1|2)\\.raw\\.fq\\.gz")
  ) %>%
  dplyr::select(`File prefix` = file_prefix, mate, md5sum) %>%
  dplyr::mutate(md5sum = md5sum %>% cell_spec(monospace = TRUE)) %>%
  tidyr::spread(mate, md5sum) %>%
  kable(caption = "MD5 checksums of raw read files") %>%
  kable_styling() %>%
  scroll_box(height = "500px")
```