---
bibliography: references.bib
csl: references.csl
output:
  html_document:
    theme: null
    df_print: kable
params:
  reports_dir: /sbidata/server/daniel/src/latest/back_end/reports/
  project_dir: /sbidata/server/daniel/userdat/example/
  db_dir: /sbidata/server/daniel/db/
---

```{r, include=FALSE}
source("defaults.R")

project <-
  paste0(params$project_dir, "/input/project.json") %>%
  jsonlite::read_json(simplifyVector = TRUE)

selected_params <- danielLib::get_selected_params(project)

phylotyping_params <- project$params$phylotyping_params[[selected_params$phylotyping]]
phylotyping_dir <- paste0(params$project_dir, "/phylotyping/", stringr::str_replace_all(selected_params$phylotyping, " ", "_"), "/")
denoising_dir <- paste0(params$project_dir, "/denoising/", stringr::str_replace_all(selected_params$denoising, " ", "_"), "/")

taxa <- c("species", "genus", "family", "order", "class", "phylum")

denoised_tbl <-
  paste0(denoising_dir, "denoised.csv") %>%
  readr::read_csv() %>%
  tidyr::gather(denoised, abundance, -sample_id)

phylotyped_tbl <-
  paste0(phylotyping_dir, "phylotyped.csv") %>%
  readr::read_csv()

classified_tbl <-
  phylotyped_tbl %>%
  dplyr::select(-kingdom) %>%
  tidyr::gather(taxrank, taxon, -sequence, -confidence) %>%
  dplyr::mutate(classified = ! is.na(taxon)) %>%
  dplyr::group_by(taxrank, classified) %>%
  dplyr::count() %>%
  tidyr::spread(classified, n) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(frac_classified = `TRUE` / (`TRUE` + `FALSE`)) %>%
  dplyr::select(taxrank, frac_classified) %>%
  dplyr::filter(! is.na(frac_classified)) %>%
  dplyr::arrange(-dplyr::row_number()) %>%
  dplyr::mutate(taxrank = taxrank %>% factor(levels = taxa, ordered = TRUE))

majority_classified_text <-
  classified_tbl %>%
  dplyr::filter(frac_classified > 0.5) %>%
  pull(taxrank) %>% {
    ifelse(
      test = purrr::is_empty(.),
      yes = "The majority of the representative sequences were not annotated up to the phylum rank",
      no = sprintf("The majority of the representative sequences were annotated at least up to %s rank", .))
    }

db <-
  paste0(params$db_dir, "/reference_db/db.csv") %>%
  readr::read_csv() %>%
  dplyr::filter(id == phylotyping_params$ref_database) %>%
  as.list()
```

# Phylotyping
Phylotyping is the process of taxonomic annotation of the denoised representative sequences.

## Methods
```{rmd, eval = phylotyping_params$sequence_classifier == "qiime2_nb"}
- Representative sequences were classified using QIIME2 Naive Bayes Classifier [@pmid31399723]
```
```{rmd, eval = phylotyping_params$ref_database %>% stringr::str_detect("^unite")}
- The database '`r {db$name}`' was used for taxonomic classification [@pmid30371820]
```
```{rmd, eval = phylotyping_params$ref_database %>% stringr::str_detect("^thf")}
- The database '`r {db$name}`' was used for taxonomic classification [@pmid30371820]
```
```{rmd, eval = phylotyping_params$sequence_classifier == "qiime2_blast"}
- A minimum sequence identity of `r {phylotyping_params$frac_identity * 100}`% was required for taxonomic annotation [@pmid25891793]
```

## Results
- `r {classified_tbl %>% dplyr::filter(taxrank == "species") %>% pull(frac_classified) %>% magrittr::multiply_by(100) %>% round(1)}`% and `r {classified_tbl %>% dplyr::filter(taxrank == "phylum") %>% pull(frac_classified) %>% magrittr::multiply_by(100) %>% round(1)}`% of all representative sequences were successfully annotated at species and phylum level, respectively
- `r {majority_classified_text}`

```{r, fig.cap = "Fraction of representative sequences classified at any given taxonomic rank"}
classified_plt <-
  classified_tbl %>%
  dplyr::filter(! is.na(taxrank)) %>%
  dplyr::mutate(perc_classified = frac_classified * 100) %>%
  ggplot2::ggplot(aes(taxrank, perc_classified, tooltip = sprintf("%.2f", perc_classified))) +
    ggiraph::geom_bar_interactive(stat = "identity") +
    ggplot2::labs(x = "Taxonomic rank", y = "Classified (%)")
plot_girafe(classified_plt)
```


```{r, fig.cap = "Fraction of denoised abundance unclassified at any given taxonomic rank"}
abundance_classified_tbl <-
  phylotyped_tbl %>%
  dplyr::inner_join(denoised_tbl, by = c("sequence" = "denoised")) %>%
  tidyr::gather(taxrank, taxon, taxa) %>%
  # TSS normalization of abundance per sample
  dplyr::group_by(sample_id) %>%
  dplyr::mutate(abundance = abundance / sum(abundance) * 100) %>%
  dplyr::filter(is.na(taxon)) %>%
  # get farction of unclassified abundance
  dplyr::group_by(sample_id, taxrank) %>%
  dplyr::summarise(abundance_classified_perc = 100 - sum(abundance)) %>%
  dplyr::mutate(taxrank = taxrank %>% factor(levels = taxa, ordered = TRUE))

abundance_classified_plt <-
   abundance_classified_tbl %>%
   dplyr::mutate(tooltip = sprintf("Sample:%s<br>Rank:%s<br>Classified:%.2f%%", sample_id, taxrank, abundance_classified_perc)) %>%
   ggplot2::ggplot(aes(sample_id, taxrank, fill = abundance_classified_perc, tooltip = tooltip)) +
     ggiraph::geom_tile_interactive() +
     ggplot2::coord_flip() +
     ggplot2::labs(x = "Sample", y = "Taxonomic rank", fill = "Classified\ndenoised\nabundance (%)")
 plot_girafe(abundance_classified_plt)
```

```{r, fig.cap = "Distribution of classification confidence among all classifications"}
phylotyped_plt <-
  phylotyped_tbl %>%
  ggplot(aes(confidence, tooltip = confidence)) +
    ggiraph::geom_histogram_interactive()
plot_girafe(phylotyped_plt)
```

## References