---
bibliography: references.bib
csl: references.csl
output:
  html_document:
    theme: null
    df_print: kable
params:
  reports_dir: /sbidata/server/daniel/latest/repo/back_end/reports/
  project_dir: /sbidata/server/daniel/latest/userdat/example_soil2
  db_dir: /sbidata/server/daniel/latest/db
---

```{r, include=FALSE}
source("defaults.R")

project <-
  paste0(params$project_dir, "/input/project.json") %>%
  jsonlite::read_json(simplifyVector = TRUE)

selected_params <- danielLib::get_selected_params(project)

analysis_params <- project$params$analysis_params[[selected_params$analysis]]
analysis_dir <- paste0(params$project_dir, "/analysis/", stringr::str_replace_all(selected_params$analysis, " ", "_"), "/")
statistics_dir <- paste0(analysis_dir, "statistics/")

pval_max <- analysis_params$max_pvalue

#' @param filter_q_value also filter for q value
filter_significant <- function(stat_tbl, pval_max = 0.05, filter_q_value = FALSE) {
  if(filter_q_value) {
    stat_tbl %>%
      dplyr::filter(q.value <= pval_max)
  } else {
    stat_tbl %>%
      dplyr::filter(p.value <= pval_max)
  }
}

any_significant <- function(stat_tbl, pval_max = 0.05, filter_q_value = FALSE) {
  stat_tbl %>%
    filter_significant(pval_max = pval_max, filter_q_value = filter_q_value) %>%
    dim() %>%
    purrr::pluck(1) %>%
    magrittr::is_greater_than(0)
}

significant_features <- function(stat_tbl) {
  stat_tbl %>%
    filter_significant() %>%
    dplyr::pull(feature) %>%
    unique() %>%
    paste(collapse = ",")
}

significant_groups <- function(stat_tbl) {
  stat_tbl %>%
    filter_significant() %>%
    dplyr::pull(group_var) %>%
    unique() %>%
    paste(collapse = ",")
}

significant_tests_count_str <- function(stat_tbl) {
  stat_tbl %>%
    filter_significant() %>%
    dim() %>%
    purrr::pluck(1) %>%
    as.character()
}

stat_tbl <-
  paste0(statistics_dir, "stat.csv") %>%
  readr::read_csv() %>%
  dplyr::arrange(-abs(effsize_value))

stat_bivariate_tbl <-
  stat_tbl %>%
  dplyr::filter(test_type == "two groups") 
stat_bivariate_test <-
  stat_bivariate_tbl$method %>%
  unique() %>%
  setdiff(NA) %>%
  purrr::pluck(1)

stat_aov_pre_tbl <-
  stat_tbl %>%
  dplyr::filter(test_type == "pre one way AOV") 
stat_aov_test <- 
  stat_aov_pre_tbl$method %>%
  unique() %>%
  setdiff(NA) %>%
  purrr::pluck(1)

stat_aov_post_tbl <- 
  stat_tbl %>%
  dplyr::filter(test_type == "post one way AOV") 

stat_cor_tbl <-
  stat_tbl %>%
  dplyr::filter(test_type == "correlation") 
```

<script>MathJax.Hub.Queue(["Typeset", MathJax.Hub]);</script>

# Statistics
Here, statistics refers to comparisons of feature abundance between sample groups.

```{rmd, eval=stat_tbl %>% nrow() == 0}
## Results

- No test was performed due to lacking sample meta data
```
```{r, eval=stat_tbl %>% nrow() == 0}
knitr::knit_exit()
```

## Methods
- A uniform threshold to reject the null hypothesis of $p \leq  `r {pval_max}`$ was chosen
- P values were adjusted for false discovery rate (FDR) in testing of multiple sample groupings. No adjustment was performed for multiple features.

```{rmd, eval = stat_tbl %>% filter(paired) %>% dim() %>% purrr::pluck(1) > 0}
- Paired difference tests were applied as indicated
- Column pair_id of sample meta data table was used to pair the samples
```

```{rmd, eval = stat_bivariate_tbl %>% dim() %>% purrr::pluck(1) > 0}
- Two means tests
  - Two means tests was applied on binary sample groups
  - Test: `r {stat_bivariate_test}`
  - Sample groups: `r {stat_bivariate_tbl$group_var %>% unique() %>% paste(collapse = ", ")}`
```

```{rmd, eval = stat_aov_pre_tbl %>% dim() %>% purrr::pluck(1) > 0}
- Three and more means tests
  - Three and more means statistics was applied on sample groups having more than two distinct nominal values
  - Test: `r {stat_aov_test}`
  - Post hoc test: `r {stat_aov_post_tbl$method %>% unique() %>% setdiff(NA) %>% purrr::pluck(1)}`
  - Sample groups: `r {stat_aov_pre_tbl$group_var %>% unique() %>% paste(collapse = ", ")}`
```

```{rmd, eval = stat_cor_tbl %>% dim() %>% purrr::pluck(1) > 0}
- Correlation tests
  - Correlation tests were applied on continuous sample meta data features
  - Sample attributes tested: `r {stat_cor_tbl$group_var %>% unique() %>% paste(collapse = ", ")}`
```

## Results

```{rmd, eval = stat_bivariate_tbl %>% any_significant()}
- Two means tests
  - Significant tests: `r {stat_bivariate_tbl %>% significant_tests_count_str()}`
  - Significant sample groups: `r {stat_bivariate_tbl %>% significant_groups()}`
  - Significant features: `r {stat_bivariate_tbl %>% significant_features()}`
```
```{rmd, eval = ! stat_bivariate_tbl %>% any_significant()}
- No two means test yielded significance
```
```{rmd, eval = stat_aov_post_tbl %>% any_significant()}
- Three and more means tests
  - Significance was evaluated on post hoc test level
  - Significant tests: `r {stat_aov_post_tbl %>% significant_tests_count_str()}`
  - Significant sample groups: `r {stat_aov_post_tbl %>% significant_groups()}`
  - Significant features: `r {stat_aov_post_tbl %>% significant_features()}`
```
```{rmd, eval = ! stat_aov_post_tbl %>% any_significant() && stat_aov_pre_tbl %>% any_significant()}
- Three and more means tests
  - No post hoc three and more means test was significant
  - Therefore, significance was evaluated on the level of the first test
  - Significant tests: `r {stat_aov_pre_tbl %>% significant_tests_count_str()`
  - Significant sample groups: `r {stat_aov_pre_tbl %>% significant_groups()}`
  - Significant features: `r {stat_aov_pre_tbl %>% significant_features()}`
```
```{rmd, eval = ! (stat_aov_post_tbl %>% any_significant() && stat_aov_pre_tbl %>% any_significant()) }
- Neither three and more means tests nor post hoc tests yielded any significant result
```
```{rmd, eval = FALSE && stat_cor_tbl %>% any_significant()}
- Correlation tests
  - Significance was evaluated on post hoc test level
  - Number of significant tests: `r {stat_cor_tbl %>% filter_significant() %>% dim() %>% pluck(1) %>% as.character()}`
  - Significant sample groups: `r {stat_cor_tbl %>% filter_significant() %>% dplyr::pull(group_var) %>% unique()}`
  - Significant features: `r {stat_cor_tbl %>% filter_significant() %>% dplyr::pull(feature) %>% unique() %>% paste(collapse = ", ")}`
```
```{rmd, eval = ! stat_cor_tbl %>% any_significant()}
- No correlation test yielded significance
```

```{r, eval = stat_bivariate_tbl %>% filter_significant() %>% nrow() > 0}
stat_bivariate_tbl %>%
  filter_significant() %>%
  dplyr::mutate(
    group_var = group_var %>%
      kableExtra::cell_spec(
        format = "HTML",
        tooltip = sample_ids %>% str_replace_all(",", ", ") %>% paste0("Samples used:", .) %>% paste0(group_counts, "\n", .)
      )
  ) %>%
  dplyr::select(
    Grouping = `group_var`, Feature = feature,
    `Paired test` = paired, `p value` = p.value, `q value` = q.value,
    effsize_unit, effsize_value, `higher in` = higher.in
  ) %>%
  tidyr::pivot_wider(names_from = effsize_unit, values_from = effsize_value) %>%
  kable(caption = "Significant two means comparisons") %>%
  collapse_rows(columns = 1:2)
```

```{r, eval = stat_aov_pre_tbl %>% filter_significant() %>% nrow() > 0}
stat_aov_pre_tbl %>%
  filter_significant() %>%
  dplyr::arrange(feature) %>%
  select(
    Grouping = `group_var`, Feature = feature,
    `p value` = p.value, `q value` = q.value
  ) %>%
  kable(caption = "Significant three or more means pre hoc tests") %>%
  collapse_rows(columns = 1:2)
```

```{r, eval = stat_aov_post_tbl %>% filter_significant() %>% nrow() > 0}
stat_aov_post_tbl %>%
  filter_significant() %>%
  dplyr::arrange(feature) %>%
  dplyr::select(
    Grouping = `group_var`, Feature = feature, `Group A` = group_a, `Group B` = group_b, `higher in` = higher.in,
    `p value` = p.value, `q value` = q.value, `Cliff's Delta`, `Log2 FC`
  ) %>%
  kable(caption = "Significant three or more means post hoc tests") %>%
  collapse_rows(columns = 1:2)
```

```{r, eval = stat_cor_tbl %>% filter_significant() %>% nrow() > 0}
stat_cor_tbl %>%
  filter_significant() %>%
  dplyr::arrange(feature) %>%
    dplyr::select(
    Grouping = `group_var`, Feature = feature,
    `p value` = p.value, `q value` = q.value,
  ) %>%
  kable(caption = "Significant correlation tests") %>%
  collapse_rows(columns = 1:2)
```

```{r, eval = "error" %in% colnames(stat_tbl) && stat_tbl %>% dplyr::filter(! is.na(error)) %>% nrow() > 0}
stat_tbl %>%
  dplyr::filter(! is.na(error)) %>%
  dplyr::group_by(error) %>%
  dplyr::summarise(
    group_vars = group_var %>% unique() %>% paste(collapse = ", "),
    features = feature %>% unique() %>% head(5) %>% paste(collapse = ", ")
  ) %>%
  dplyr::rename(`Grouping` = group_vars, Error = error, `Features (max. 5)` = features) %>%
  kable(caption = "Errors occurred in statistical tests")
```
