---
bibliography: references.bib
csl: references.csl
output:
  html_document:
    theme: null
    df_print: kable
params:
  reports_dir: /sbidata/server/daniel/latest/repo/back_end/reports/
  project_dir: /sbidata/server/daniel/latest/userdat/example_soil
  db_dir: /sbidata/server/daniel/latest/db
---

```{r, include=FALSE}
source("defaults.R")

project <-
  paste0(params$project_dir, "/input/project.json") %>%
  jsonlite::read_json(simplifyVector = TRUE)

selected_params <- danielLib::get_selected_params(project)

analysis_dir <- paste0(params$project_dir, "/analysis/", stringr::str_replace_all(selected_params$analysis, " ", "_"), "/")
statistics_dir <- paste0(analysis_dir, "statistics/")
correlation_dir <- paste0(analysis_dir, "correlation/")

analysis_params <- project$params$analysis_params[[selected_params$analysis]]
features_params <- project$params$features_params[[selected_params$features]]

pval_max <- analysis_params$max_pvalue

findings_tbl <- tibble::tibble()
```

```{r, include=FALSE, eval = paste0(statistics_dir, "stat.csv") %>% readr::read_csv() %>% nrow() > 0}
stat_tbl <-
  paste0(statistics_dir, "stat.csv") %>%
  readr::read_csv()

stat_tbl <-
  stat_tbl %>%
  dplyr::mutate(relationship_type = "statistics") %>%
  dplyr::select(
    feature, relationship_type,
    interaction_method = method, interaction = group_var,
    effsize_unit, effsize_value, p_value = p.value
  )
findings_tbl <-
  findings_tbl %>%
  dplyr::bind_rows(stat_tbl)
```

```{r, eval=paste0(correlation_dir, "/*/results.csv") %>% Sys.glob() %>% length() > 0}
cor_tbl <-
  paste0(correlation_dir, "/*/results.csv") %>%
  Sys.glob() %>%
  purrr::map(readr::read_csv) %>%
  bind_rows() %>%
  transmute(
    feature_a = feature_a,
    feature_b = feature_b,
    effsize_unit = analysis_params$correlation_method,
    effsize_value = cor,
    p_value = p_value
  )

cor_tbl <-
  dplyr::bind_rows(
    # inflate pairwise interactions
    cor_tbl %>% rename(feature = feature_a, interaction = feature_b),
    cor_tbl %>% rename(feature = feature_b, interaction = feature_a)
  ) %>%
  dplyr::mutate(relationship_type = "correlation", interaction_method = "sparcc")

findings_tbl <-
  findings_tbl %>%
  dplyr::bind_rows(cor_tbl)
```

```{r, eval=findings_tbl %>% nrow() > 0}
findings_tbl <-
  findings_tbl %>%
  dplyr::select(feature, relationship_type, interaction_method, interaction, effsize_unit, effsize_value, p_value) %>%
  dplyr::filter(p_value <= pval_max)

#
# annotate interactions
#

findings_features <-
  findings_tbl %>%
  pull(feature) %>%
  unique() %>%
  map_chr(~ .x %>% str_remove(" sp.$"))

lineage_tbl <-
  paste0(params$db_dir, "/phylogenies/index_fungorum_2016/taxonomy.csv.gz") %>%
  readr::read_csv()

mesh_link <- function(mesh_id) {
  if (is.na(mesh_id)) {
    return("")
  }

  shiny::tags$a(href = sprintf("https://id.nlm.nih.gov/mesh/%s.html", mesh_id), "MeSH")
}

daniel_db_con <-
  paste0(params$db_dir, "/daniel_db/daniel_db.sqlite") %>%
  DBI::dbConnect(RSQLite::SQLite(), .)

interactions_tbl <-
  daniel_db_con %>%
  DBI::dbReadTable("interactions") %>%
  tibble::as_tibble() %>%
  dplyr::rename(species = fungus) %>%
  dplyr::inner_join(lineage_tbl) %>%
  dplyr::mutate(feature = !!as.name(features_params$taxonomic_rank)) %>%
  dplyr::rename(relationship_type = interaction_type) %>%
  dplyr::filter(feature %in% findings_features) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    interaction = paste(
      relationship_type %>% str_replace_na(""),
      "interaction between",
      shiny::tags$i(species),
      "and",
      partner %>% stringr::str_replace_na(""),
      comment %>% stringr::str_replace_na(""),
      paste0(
      "(",
      shiny::tags$a(href = paste0("https://www.ncbi.nlm.nih.gov/pubmed/?term=", pubmed_id), "Reference"),
      ifelse(is.na(partner_mesh_id), "", paste0(", ", partner_mesh_id %>% mesh_link())),
      ")" 
      )
    ),
    relationship_type = "interactions_db"
  ) %>%
  dplyr::select(feature, relationship_type, interaction)

infections_tbl <-
  daniel_db_con %>%
  DBI::dbReadTable("infections") %>%
  tibble::as_tibble() %>%
  dplyr::inner_join(lineage_tbl, by = "species") %>%
  dplyr::mutate(feature = !!as.name(features_params$taxonomic_rank)) %>%
  dplyr::filter(feature %in% findings_features) %>%
  dplyr::transmute(
    feature = feature,
    relationship_type = "infections_db",
    interaction = sprintf("<i>%s</i> was found in %s", species, source)
  )

funguild_features <- c("guild", "trophicMode", "growthForm", "confidenceRanking", "url", "citationSource")
funguild_tbl <-
  base::paste0(params$db_dir, "/etc/FUNGuild.csv") %>%
  readr::read_csv() %>%
  dplyr::distinct_at(c("taxon", features_params$taxonomic_rank, funguild_features)) %>%
  #tidyr::pivot_longer(cols = funguild_features) %>%
  dplyr::rename(feature = !!sym(features_params$taxonomic_rank)) %>%
  dplyr::mutate(relationship_type = "funguild") %>%
  dplyr::filter(feature %in% findings_features)

findings_tbl <-
  findings_tbl %>%
  dplyr::full_join(interactions_tbl) %>%
  dplyr::full_join(infections_tbl) %>%
  dplyr::full_join(funguild_tbl)

findings_tbl %>%
  readr::write_csv(paste0(analysis_dir, "summary/findings.csv"))
```

# Summary
Here, significant results from all tests are combined.
Furthermore, significant features were annotated with known interactions and clinical isolates.

```{rmd, eval=findings_tbl %>% nrow() == 0}
## Results

- No feature could be annotated due to lacking statsitical tests and correlation networks
```

```{r, eval=findings_tbl %>% nrow() == 0}
findings_tbl %>%
  readr::write_csv(paste0(analysis_dir, "summary/findings.csv"))
knitr::knit_exit()
```

## Methods
DAnIEL ITS pipeline was run to analyze these relationships of fungal features:

- Statistics: Fungal features significantly differentially abundant in sample groups
- Correlation networks: Fungal features significantly correlated to other fungal features
- Literature annotation: Known interactions using a manually curated literature abstract database
- Clinical annotation: Species found in clinical isolates with suspicious fungal infections in the [NRZMyk](https://www.nrz-myk.de/home.html) reference database
    
## Results
```{r}
summary_html <- function(findings_tbl, relationship_type) {
  switch(relationship_type,
    "correlation" = {
      findings_tbl %>%
        dplyr::mutate(interaction = sprintf("<i>%s</i>", interaction)) %>%
        dplyr::mutate(relationship_html = sprintf("<li><i>%s</i> (%s, p=%.2e)</li>", interaction, effsize_unit, p_value)) %>%
        dplyr::pull(relationship_html) %>%
        paste(collapse = " ") %>%
        paste0("Significantly correlated with: <br> <ul>", ., "</ul>")
    },
    "statistics" = {
      findings_tbl %>%
        dplyr::mutate(relationship_html = sprintf("<li>%s (%s, p=%.2e)</li>", interaction, interaction_method, p_value)) %>%
        dplyr::pull(relationship_html) %>%
        paste(collapse = " ") %>%
        paste0("Significantly differentially abundant in: <br> <ul>", ., "</ul>")
    },
    "interactions_db" = {
      findings_tbl %>%
        dplyr::mutate(relationship_html = sprintf("<li>%s</li>", interaction)) %>%
        dplyr::pull(relationship_html) %>%
        paste(collapse = " ") %>%
        paste0("Known interactions from literature: <br> <ul>", ., "</ul>")
    },
    "infections_db" = {
      findings_tbl %>%
        dplyr::mutate(relationship_html = sprintf("<li>%s</li>", interaction)) %>%
        dplyr::pull(relationship_html) %>%
        paste(collapse = " ") %>%
        paste0("Clinical isolates with suspicious fungal infections: <br> <ul>", ., "</ul>")
    },
    "funguild" = {
      findings_tbl %>%
        dplyr::mutate(reference = case_when(
          ! is.na(url) ~ stringr::str_glue("<a href='{url}' title='{citationSource}'>Reference</a>"),
          ! is.na(citationSource) ~ citationSource,
          TRUE ~ ""
        )) %>%
        dplyr::mutate(relationship_html = stringr::str_glue(
          "<li>{growthForm %>% replace_na('')} <i>{taxon}</i> is of guild {guild} ",
          "and trophic mode {trophicMode} ({confidenceRanking}, {reference})"
        )) %>%
        dplyr::pull(relationship_html) %>%
        paste(collapse = " ") %>%
        paste0("Functions (FUNGuild): <br> <ul>", ., "</ul>")
    },
    paste0(relationship_type, " not yet implemented.")
  )
}
  
findings_summary_tbl <-
  findings_tbl %>%
  dplyr::group_by(feature, relationship_type) %>%
  tidyr::nest() %>%
  dplyr::mutate(
    summary_html = purrr::pmap_chr(list(data, relationship_type), summary_html)
  )

findings_summary_tbl %>%
  dplyr::group_by(feature) %>%
  dplyr::summarise(
    feature_summary_html = paste(summary_html, collapse = "<br>") %>% paste(sprintf("<h3><i>%s</i></h3>", feature %>% first()), .)
  ) %>%
  pull(feature_summary_html) %>%
  shiny::HTML()
```
