---
bibliography: "/sbidata/server/daniel/latest/db/etc/daniel.bib"
csl: "/sbidata/server/daniel/latest/db/etc/daniel.csl" 
output:
  html_document:
    theme: null
---

```{r setup, include=FALSE}
library(tidyverse)
library(kableExtra)
library(knitr)
library(rmarkdown)
library(danielLib)

danielLib::setup_style()
options(knitr.kable.NA = "")

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

steps <- c("input", "qc", "denoising", "phylotyping", "features", "analysis")
```

# Reference

This page is about the specification of the server.
A detailed description about tools, parameters, and tested browsers is given.

## Tools
The general framework is based on Snakemake [@pmid29788404], R, Tidyverse, and Docker.

```{r}
tools_tbl <-
  Sys.getenv("DANIEL_DB_DIR") %>%
  paste0("/etc/tools.csv") %>%
  readr::read_csv() %>%
  dplyr::mutate(param_set = param_set %>% factor(levels = steps, ordered = TRUE))

tools_tbl %>%
  arrange(param_set, id) %>%
  select(`Pipeline step` = param_set, Tool = id, Description = description, Version = version, Citation = citation, Source = source) %>%
  kable(caption = "Tools") %>%
  column_spec(3, width = "20em") %>%
  kableExtra::collapse_rows(1)
```

## Parameters

In the input and start section, the following parameters can be set to customize the workflow:

```{r}
param_fmt_html <- "
<p>
  <p>
    <b>%s</b> <br />
    %s <br />
  </p>
  <p>
    Type: %s <br />
    Default: %s <br />
  </p>
</p>
"

#' @param sep (sub)string separating two elements of s
#' @param s string representing a list. Elements must be separated by sep
#' @return s if s is single element, else HTML list of s
html_if_needed <- function(s, sep = "~") {
  if(is.na(s)) return(NA)
  if(! stringr::str_detect(s, sep)) return(s)
  
  s %>%
    str_split("~") %>%
    simplify() %>%
    map(~ paste0("<li>", .x, "</li>")) %>%
    paste0(collapse = " ") %>%
    paste0("<ul>", ., "</ul>")
}

params_tbl <-
  Sys.getenv("DANIEL_DB_DIR") %>%
  paste0("/etc/params.csv") %>%
  readr::read_csv() %>%
  dplyr::mutate(param_set = param_set %>% factor(levels = steps, ordered = TRUE))

params_tbl %>%
  dplyr::group_by(id) %>%
  dplyr::mutate(
    default_html = default %>% html_if_needed(),
    html = sprintf(param_fmt_html, label, description, type, default_html)
  ) %>%
  purrr::pluck("html") %>%
  shiny::HTML()
```

## Browsers

DAnIEL was sucessfully tested with the following browers:

```{r}
browsers_tbl <-
  Sys.getenv("DANIEL_DB_DIR") %>%
  paste0("/etc/browsers.csv") %>%
  readr::read_csv(col_types = cols(Versions = col_character()))

browsers_tbl %>% 
  mutate(Versions = as.character(Versions)) %>%
  kable(caption = "Browsers")
```


## References
